On the properties of projectile motion and quadratic air resistance

### 2016-04-04 Paper/main.tex
Minor align
Printed
Fixed page numbering
Minor

\setcounter{page}{4}

@@

		\end{algorithmic}
	\end{algorithm}
	Using the above algorithm with initial data $v_0 =10,m/s$, $\theta = \ang{45} = \frac{\pi}{4}$, $\rho = 1.225\, kg/m^3$, $A = 1\,m^2$, $m = 1\,kg$, $y_0 = 0\,m$ and $C_D = 0.47$ yields the following result. The solid plot is with air resistance and the dashed plot is in a vacuum with the same $v_0$, $y_0$ and $\theta$.
	\begin{center}
		\begin{tikzpicture}
		\begin{axis}[ 
		xlabel=x (m),
		ylabel=y (m)
		] 
		\addplot[solid] coordinates {(0.0000, 0.0000) (0.2674, 0.2597) (0.5092, 0.4794) (0.7305, 0.6654) (0.9349, 0.8222) (1.1255, 0.9531) (1.3043, 1.0606) (1.4732, 1.1469) (1.6335, 1.2135) (1.7862, 1.2616) (1.9323, 1.2922) (2.0724, 1.3061) (2.2071, 1.3041) (2.3367, 1.2868) (2.4615, 1.2546) (2.5818, 1.2082) (2.6977, 1.1481) (2.8093, 1.0747) (2.9166, 0.9888) (3.0197, 0.8909) (3.1185, 0.7815) (3.2132, 0.6614) (3.3037, 0.5312) (3.3902, 0.3916) (3.4725, 0.2432) (3.5509, 0.0866) (3.5923, -0.0028) };

@@

		y''(t) = k\,y'(t)^2 - g \comma k = \frac{\frac12\rho\,A\,C_D}{m}\\\nonumber
		y(0) = y_0 \comma y'(0) = 0

@@

		\addplot[solid, dashed] coordinates {(0.000, 4.000) (0.010, 4.000) (0.020, 3.998) (0.030, 3.996) (0.040, 3.992) (0.050, 3.988) (0.060, 3.982) (0.070, 3.976) (0.080, 3.969) (0.090, 3.960) (0.100, 3.951) (0.110, 3.941) (0.120, 3.929) (0.130, 3.917) (0.140, 3.904) (0.150, 3.890) (0.160, 3.874) (0.170, 3.858) (0.180, 3.841) (0.190, 3.823) (0.200, 3.804) (0.210, 3.783) (0.220, 3.762) (0.230, 3.740) (0.240, 3.717) (0.250, 3.693) (0.260, 3.668) (0.270, 3.642) (0.280, 3.615) (0.290, 3.587) (0.300, 3.558) (0.310, 3.528) (0.320, 3.497) (0.330, 3.465) (0.340, 3.432) (0.350, 3.399) (0.360, 3.364) (0.370, 3.328) (0.380, 3.291) (0.390, 3.253) (0.400, 3.214) (0.410, 3.175) (0.420, 3.134) (0.430, 3.092) (0.440, 3.049) (0.450, 3.006) (0.460, 2.961) (0.470, 2.915) (0.480, 2.869) (0.490, 2.821) (0.500, 2.772) (0.510, 2.723) (0.520, 2.672) (0.530, 2.621) (0.540, 2.568) (0.550, 2.515) (0.560, 2.460) (0.570, 2.405) (0.580, 2.348) (0.590, 2.291) (0.600, 2.232) (0.610, 2.173) (0.620, 2.113) (0.630, 2.051) (0.640, 1.989) (0.650, 1.926) (0.660, 1.861) (0.670, 1.796) (0.680, 1.730) (0.690, 1.662) (0.700, 1.594) (0.710, 1.525) (0.720, 1.455) (0.730, 1.383) (0.740, 1.311) (0.750, 1.238) (0.760, 1.164) (0.770, 1.089) (0.780, 1.013) (0.790, 0.936) (0.800, 0.858) (0.810, 0.779) (0.820, 0.699) (0.830, 0.618) (0.840, 0.536) (0.850, 0.453) (0.860, 0.369) (0.870, 0.284) (0.880, 0.198) (0.890, 0.111) (0.900, 0.023) (0.910, -0.066) };
		\end{axis}
		\end{tikzpicture}
	\end{center}
	The slow start with a negative second derivative matches our intuition and at roughly $t = 0.8\,s$ the first derivative appears to approach zero. This should be due to the object reaching its terminal velocity which we can calculate using equation \eqref{eq:oneDimAirResTerminalVel}.
	\begin{align}\nonumber
		v_{max} = \sqrt{\frac{g}{k}} = \sqrt{\frac{9.82}{0.432}} = 4.77\,m/s
	\end{align}
	We can approximate the derivative at $t=0.8$ as $\frac{\Delta y}{\Delta t} \approx 4.4 \approx v_{max}$. Equation \eqref{eq:oneDimAirResTMax} also predicts a $t_{max}$ which can be tested.
	\begin{align}\nonumber
		t_{max} = \frac{\cosh^{-1}(\exp(y_0\,k))}{\sqrt{gk}} = \frac{\cosh^{-1}(\exp(1.728))}{\sqrt{4.424224}} \approx 1.17\,s
	\end{align}
	The plot reports $t_{max} \approx 1.17\,s$. Comparing this to a vacuum solution we can use equation \eqref{eq:tMaxFromHeight} which gives
	\begin{align}\nonumber
		t_{max} &= \frac{v_0\sin\theta}{g}+\frac{\sqrt{v_{0}^2\sin^2\theta+2g\,y_0}}{g}\\\nonumber
		&= \frac{\sqrt{2\,g\,y_0}}{g} = \frac{\sqrt{78.56}}{9.82} \approx 0.90\,s
	\end{align}
	This is a major discrepancy which adds credit to the accuracy of equation \eqref{eq:oneDimAirResTMax}.

>>>>

\setcounter{page}{1}

@@

		\end{algorithmic}
	\end{algorithm}
	Using the above algorithm with initial data $v_0 =10,m/s$, $\theta = \ang{45} = \frac{\pi}{4}$, $\rho = 1.225\, kg/m^3$, $A = 1\,m^2$, $m = 1\,kg$, $y_0 = 0\,m$ and $C_D = 0.47$ which corresponds to a heavy beach ball yields the following result. The solid plot is with air resistance and the dashed plot is in a vacuum with the same $v_0$, $y_0$ and $\theta$.
	\begin{center}
		\begin{tikzpicture}
		\begin{axis}[ 
		xlabel=x (m),
		ylabel=y (m)
		] 
		\addplot[solid] coordinates {(0.0000, 0.0000) (0.2674, 0.2597) (0.5092, 0.4794) (0.7305, 0.6654) (0.9349, 0.8222) (1.1255, 0.9531) (1.3043, 1.0606) (1.4732, 1.1469) (1.6335, 1.2135) (1.7862, 1.2616) (1.9323, 1.2922) (2.0724, 1.3061) (2.2071, 1.3041) (2.3367, 1.2868) (2.4615, 1.2546) (2.5818, 1.2082) (2.6977, 1.1481) (2.8093, 1.0747) (2.9166, 0.9888) (3.0197, 0.8909) (3.1185, 0.7815) (3.2132, 0.6614) (3.3037, 0.5312) (3.3902, 0.3916) (3.4725, 0.2432) (3.5509, 0.0866) (3.5923, -0.0028) };

@@

		y''(t) &= k\,y'(t)^2 - g \comma k = \frac{\frac12\rho\,A\,C_D}{m}\\\nonumber
		y(0) &= y_0 \comma y'(0) = 0

@@

		\addplot[solid, dashed] coordinates {(0.000, 4.000) (0.010, 4.000) (0.020, 3.998) (0.030, 3.996) (0.040, 3.992) (0.050, 3.988) (0.060, 3.982) (0.070, 3.976) (0.080, 3.969) (0.090, 3.960) (0.100, 3.951) (0.110, 3.941) (0.120, 3.929) (0.130, 3.917) (0.140, 3.904) (0.150, 3.890) (0.160, 3.874) (0.170, 3.858) (0.180, 3.841) (0.190, 3.823) (0.200, 3.804) (0.210, 3.783) (0.220, 3.762) (0.230, 3.740) (0.240, 3.717) (0.250, 3.693) (0.260, 3.668) (0.270, 3.642) (0.280, 3.615) (0.290, 3.587) (0.300, 3.558) (0.310, 3.528) (0.320, 3.497) (0.330, 3.465) (0.340, 3.432) (0.350, 3.399) (0.360, 3.364) (0.370, 3.328) (0.380, 3.291) (0.390, 3.253) (0.400, 3.214) (0.410, 3.175) (0.420, 3.134) (0.430, 3.092) (0.440, 3.049) (0.450, 3.006) (0.460, 2.961) (0.470, 2.915) (0.480, 2.869) (0.490, 2.821) (0.500, 2.772) (0.510, 2.723) (0.520, 2.672) (0.530, 2.621) (0.540, 2.568) (0.550, 2.515) (0.560, 2.460) (0.570, 2.405) (0.580, 2.348) (0.590, 2.291) (0.600, 2.232) (0.610, 2.173) (0.620, 2.113) (0.630, 2.051) (0.640, 1.989) (0.650, 1.926) (0.660, 1.861) (0.670, 1.796) (0.680, 1.730) (0.690, 1.662) (0.700, 1.594) (0.710, 1.525) (0.720, 1.455) (0.730, 1.383) (0.740, 1.311) (0.750, 1.238) (0.760, 1.164) (0.770, 1.089) (0.780, 1.013) (0.790, 0.936) (0.800, 0.858) (0.810, 0.779) (0.820, 0.699) (0.830, 0.618) (0.840, 0.536) (0.850, 0.453) (0.860, 0.369) (0.870, 0.284) (0.880, 0.198) (0.890, 0.111) (0.900, 0.023) (0.910, -0.066) };
		\end{axis}
		\end{tikzpicture}
	\end{center}
	The slow start with a negative second derivative matches our intuition and at roughly $t = 0.8\,s$ the second derivative appears to approach zero. This should be due to the object reaching its terminal velocity which we can calculate using equation \eqref{eq:oneDimAirResTerminalVel}.
	\begin{align}\nonumber
		v_{max} = \sqrt{\frac{g}{k}} = \sqrt{\frac{9.82}{0.432}} = 4.77\,m/s
	\end{align}
	We can approximate the derivative at $t=0.8$ as $\frac{\Delta y}{\Delta t} \approx 4.4 \approx v_{max}$. Equation \eqref{eq:oneDimAirResTMax} also predicts a $t_{max}$ which can be tested.
	\begin{align}\nonumber
		t_{max} = \frac{\cosh^{-1}(\exp(y_0\,k))}{\sqrt{gk}} = \frac{\cosh^{-1}(\exp(1.728))}{\sqrt{4.424224}} \approx 1.17\,s
	\end{align}
	The plot reports $t_{max} \approx 1.17\,s$. Comparing this to a vacuum solution we can use equation \eqref{eq:tMaxFromHeight} which gives
	\begin{align}\nonumber
		t_{max} &= \frac{v_0\sin\theta}{g}+\frac{\sqrt{v_{0}^2\sin^2\theta+2g\,y_0}}{g}\\\nonumber
		&= \frac{\sqrt{2\,g\,y_0}}{g} = \frac{\sqrt{78.56}}{9.82} \approx 0.90\,s
	\end{align}
	This is a major discrepancy which adds credit to the accuracy of equation \eqref{eq:oneDimAirResTMax}.
